---
title: "hw01-Betty-Cui.Rmd"
author: "Betty Cui"
date: "2/18/2018"
output: github_document
---

### Data Dictionary (To build a data dictionary in a separated text file) 
cd Desktop/stats133  
mkdir hw01  
cd hw01  
mkdir imports-85-dictionary.md  

```{r}
## Data Import
column_names <- c("symboling","normalized_losses","make","fuel_type","aspiration","num_of_doors","body_style","drive_wheel", "engine_location", "wheel_base", "length","width","height", "curb_weight", "engine_type", "num_of_cylinders", "engine_size", "fuel_system", "bore", "stroke", "compression_ratio", "horsepower", "peak_rpm", "city_mpg", "highway_mpg", "price")

classes <- c( "real","character","character", "character", "character", "character","character", "character", "character", "real","real","real", "real", "integer", "character", "character", "integer", "character", "real", "real", "real", "integer", "integer", "integer","integer", "integer")


library(readr)
dat <- data.frame(
  read_csv("./imports-85.data",
           col_types = cols(symboling = col_double(),
                            normalized.losses = col_character(), 
                            make = col_character(), 
                            fuel_type = col_character(), 
                            aspiration = col_character(), 
                            num_of_doors = col_character(), 
                            body_style = col_character(), 
                            drive_wheel = col_character(), 
                            engine_location = col_character(), 
                            wheel_base = col_double(), 
                            length = col_double(), 
                            width = col_double(), 
                            height = col_double(), 
                            curb_weight = col_integer(), 
                            engine_type = col_character(), 
                            num_of_cylinders = col_character(), 
                            engine_size = col_integer(), 
                            fuel_system = col_character(), 
                            bore = col_double(), 
                            stroke = col_double(), 
                            compression_ratio = col_double(), 
                            horsepower = col_integer(), 
                            peak_rpm = col_integer(), 
                            city_mpg = col_integer(), 
                            highway_mpg = col_integer(), 
                            price = col_integer())
           )
  )

str(dat)


dat <- read.csv("imports-85.data",header = FALSE, col.names = column_names, colClasses = classes, na.strings="?",  stringsAsFactors = FALSE)
str(dat)


```


### 3 Technical Questions about importing data
a) The column name is used as column names and the real data values starts from the second line.
b) Since the header is set to False, the first line(column names) will be considered as data value.
c) Each data type will be factor.
d) Because on the first option, we specify the column type which requires more storage. 
e) It is going to convert to three columns, one column is value, one is variable name, and another one is data(index)

```{r}
## Colored Historgram of Price
colors <- c("yellow", "red", "black", "green", "violet", "orange", "blue", "pink", "cyan")
hist(dat$price, col = colors)

## Boxplot of horsepower in horizontal orientation
horsepower <- dat$horsepower
boxplot(horsepower, horizontal = TRUE)

## Barplot of frequencies if body_style in decreasing order
body_style <- dat$body_style
body_style <- table(body_style)
barplot(sort(body_style, decreasing = TRUE))
colnames <- c("wheel_base", "length", "width", "height", "price")
turbo1 <- dat[dat$aspiration == 'turbo' , colnames]
stars(turbo1, key.labels = colnames)
```

```{r}
## Summaries
#a) Mean price of fuel_type gas car
fuel <- dat$fuel_type
total <- aggregate(dat$price, by = list(fuel), mean, na.rm = TRUE)
total
#b) make of twelve num_of_cylinders 
make <- dat$make[dat$num_of_cylinders == "twelve"]
make
#c) make that has most diesel car
make1 <- dat$make[dat$fuel_type == "diesel"]
count_frequency <- sort(table(make1), decreasing = TRUE)
new_make <- names(count_frequency[1])
new_make

#d) price of the car with largest hoursepower
price <- dat$price[dat$horsepower]
price
largest_horsepower <- sort(table(price),decreasing = TRUE)
largest_horsepower
new_price <- names(largest_horsepower[1])
new_price
#e) botthom 10th percentile of city_mpg
a <- quantile(dat$city_mpg, 0.1, na.rm=TRUE)
a
#f) top 10th percentile of highway_mpg
b <- quantile(dat$highway_mpg, 0.9, na.rm=TRUE)
b
#g) median_price of botthom 10th percentile of city_mpg
c <- median(dat$price[a])
c
as.list(dat)
```

###Technical Questions
* If to use the name that does not exist, then error and shows column type is NUll.
* 4 fails because object 'mpg', which is not a string, can not be found
* Yes, A data.frame is a list of vectors, each with the same length and a list is a type of vector.
* The whole data appears in the form of lists
* Use data.frame(abc)
* Each variable and its values are converted to individual list.
```{r}

#Correlations of quantitative variables
quantative_data <- dat[ , c("symboling", "wheel_base", "length","width","height", "curb_weight", "engine_size", "bore", "stroke", "compression_ratio", "horsepower", "peak_rpm", "city_mpg", "highway_mpg", "price")]
qdat <- na.omit(quantative_data)
qdat <- cor(qdat)
head(round(qdat, 2))
library(corrplot)
corrplot(qdat, method = "circle")
corrplot.mixed(qdat, lower = "number", upper = "shade", lower.col = "black", number.cex = 0.7)
corrplot(qdat, type = "upper", order = "hclust", col = c("black", "white"), bg ="lightblue")
corrplot(qdat,type = "upper", order = "hclust",tl.col = "black", tl.srt = 45)
cor.dat <- function(mat,... ){
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n-1)){
    for (j in (i+1):n){
      tmp <- cor.test(mat[,i], mat[, j],... )
      p.mat[i,j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
p.mat <- cor.dat(qdat)
head(p.mat[, 1:5])
corrplot(qdat,type = "upper", order = "hclust", p.mat = p.mat, sig.level = 0.01)
```


```{r}
#8 Principal Components Analysis
# 8.1 
pca_prcomp <- prcomp(qdat, scale.= TRUE)
pca_prcomp
eigenvalues <- pca_prcomp$sdev^2
eigenvalues
sum(eigenvalues[1:3])/sum(eigenvalues)
# Linear correlation between eigenvalue and percentage
```
8.2) PCA plot of vehicles, and PCA plot of variables

*
```{r}
t<-pca_prcomp$x
plot(t[,1:2])
abline(h=0, v=0)
```

*
```{r}
rot<-pca_prcomp$rotation
rot
plot(rot[,1],rot[,2])
abline(h=0, v=0)
```

*
```{r}
biplot(pca_prcomp, scale = 0)
```

*
```{r}
##  if the arrow points to the similiar direction (angle less than 90), the correlation coefficient will be positive, otherwise, else if angle ranges in 90-180, the correlation coefficient will be negative.
```




