---
title: "lab10-Betty-Cui"
author: "Betty Cui"
date: "4/4/2018"
output: html_document
---
## lab10-regualr expression
```{r}
## Data setup
library(dplyr)
library(stringr)
library(plotly)
github <- "https://raw.githubusercontent.com/ucb-stat133/stat133-spring-2018/master/"
datafile <- "data/mobile-food-sf.csv"
download.file(paste0(github, datafile), destfile = "mobile-food-sf.csv")
dat <- read.csv('mobile-food-sf.csv', stringsAsFactors = FALSE)
```

```{r}
## Your Turn: What about times where the hour has just one digit? For example: 9AM, or 8AM? Create the following vector times and try to subset the hour and the periods with str_sub()
times <- c('12PM', '10AM', '9AM', '8AM')
str_sub(times, start = -2)
str_sub(times, start = 1, end = nchar(times) - 2)
str_replace(times, pattern = 'AM|PM', replacement = '')
```

```{r}
## Your Turn So far you’ve managed to get the hour value and the period (AM or PM). Now: Using times, create a numeric vector hours containing just the number time (i.e. hour) 
hours <- as.numeric(str_replace(times, pattern = 'AM|PM', replacement = ''))
hours

## Using times, create a character vector periods containing the period, e.g. AM or PM
periods <- str_sub(times, start = -2)
periods


## Use plot_ly() to make a barchart of the counts for AM and PM values.
plot_ly(x = periods,
        y = hours,
        type = 'bar')
        

## Write R code to create a vector start24 that contains the hour in 24hr scale.
start24 <- rep(0, length(times))
for(i in 1:length(hours)){
  if(periods[i] == "PM"){
    start24[i] <- hours[i] + 12
  }else{
    start24[i] <- hours[i]
  }
}
start24
## Add two columns start and end to the data frame dat, containing the starting and ending hour respectively (columns must be "numeric"). 
start <- as.numeric(str_replace(dat$starttime, pattern = 'AM|PM', replacement = ''))
periods1 <- periods <- str_sub(dat$starttime, start = -2)
for(i in 1:length(start)){
  if(periods1[i] == "PM"){
    start[i] <- start[i] + 12
  }else{
    start[i] <- start[i]
  }
}
start
dat <- mutate(dat, start)

end <- as.numeric(str_replace(dat$endtime, pattern = 'AM|PM', replacement = ''))
periods2 <- periods <- str_sub(dat$endtime, start = -2)
for(i in 1:length(end)){
  if(periods2[i] == "PM"){
    end[i] <- end[i] + 12
  }else{
    end[i] <- end[i]
  }
}
end
dat <- mutate(dat, end)

## With the starting and ending hours, calculate the duration, and add one more column duration to the data frame dat:
duration <- end - start
for(i in 1:length(duration)){
  if(duration[i] < 0){
    duration[i] <- duration[i] + 24
  }
}
dat <- mutate(dat, duration)
```

```{r}
## My turn: create a list lon by using lapply() with anonymous function to extract longitude value
loc1 <- "(37.7651967350509,-122.416451692902)"
str_replace(loc1, pattern = '\\(', replacement = '')
locs <- c(
  "(37.7651967350509,-122.416451692902)",
  "(37.7907890558203,-122.402273431333)",
  "(37.7111991003088,-122.394693339395)",
  "(37.7773000262759,-122.394812784799)",
  NA
)
lat_lon <- str_replace_all(locs, pattern = '\\(|\\)', replacement = '')
lat_lon <- str_split(lat_lon, pattern = ',')

lat <- lapply(lat_lon, function(x) x[1])
lat_lon <- str_replace_all(dat$Location, pattern = '\\(|\\)', replacement = '')
lat_lon <- str_split(lat_lon, pattern = ',')

lat <- lapply(lat_lon, function(x) x[1])
lon <- lapply(lat_lon, function(x) x[2])

latitute <- as.numeric(unlist(lat))
longitude <- as.numeric(unlist(lon))
dat <-mutate(dat, latitute, longitude)
```

```{r}
# install.packages("RgoogleMaps")
library(RgoogleMaps)
```

```{r}
# coordinates for center of the map
center <- c(mean(dat$latitute, na.rm = TRUE), mean(dat$longitude, na.rm = TRUE))

# zoom value
zoom <- min(MaxZoom(range(dat$latitute, na.rm = TRUE), 
                    range(dat$longitude, na.rm = TRUE)))

# san francisco map
map1 <- GetMap(center=center, zoom=zoom, destfile = "san-francisco.png")

PlotOnStaticMap(map1, dat$latitute, dat$longitude, col = "#ed4964", pch=20)
```

Another useful package for plotting maps is "ggmap". As you may guess, "ggmap" follows the graphing approach of "ggplot2".

```{r}
library(ggmap)
```

```{r}
dat <- na.omit(dat)
```

```{r}
# ggmap typically asks you for a zoom level, 
# but we can try using ggmap's make_bbox function:
sbbox <- make_bbox(lon = dat$lon, lat = dat$lat, f = .1)
sbbox
```

```{r}
sf_map <- get_map(location = sbbox, maptype = "terrain", source = "google")
```

```{r}
ggmap(sf_map) + 
  geom_point(data = dat, 
             mapping = aes(x = longitude, y = latitute), 
             color = "red", alpha = 0.2, size = 1)
```


### Let's look for specific types of food

```{r}
dat$optionaltext[1:3]
```

```{r}
foods <- dat$optionaltext[1:10]
```

Use str_detect() (or equivalently grep()) to match "Burritos" and "burritos".

```{r}
str_detect(foods, pattern = "Burritos|burritos")
grepl(pattern = "burritos", foods, ignore.case = TRUE)
```

Try another pattern: e.g. "tacos", or "quesadillas"
```{r}
grepl(pattern = "tacos", foods, ignore.case = TRUE)
grepl(pattern = "quesadillas", foods, ignore.case = TRUE)

```

Use the lat and lon corrdinates in burritos to display a map of locations with burritos (see map below).
```{r}
burritos <- grepl(pattern = "burritos", dat$optionaltext, ignore.case = TRUE)
ggmap(sf_map) + 
  geom_point(data = dat[burritos, ], 
             mapping = aes(x = longitude, y = latitute), 
             color = "blue", alpha = 0.2, size = 1)
```

Other type of food
```{r}
quesadillas <- grepl(pattern = "quesadillas", dat$optionaltext, ignore.case = TRUE)
ggmap(sf_map) + 
  geom_point(data = dat[quesadillas, ], 
             mapping = aes(x = longitude, y = latitute), 
             color = "red", alpha = 0.2, size = 1)
```

Challenge
```{r}
tacos <- grepl(pattern = "tacos", dat$optionaltext, ignore.case = TRUE)
ggmap(sf_map) + 
  geom_point(data = dat[tacos, ], 
             mapping = aes(x = longitude, y = latitute), 
             color = "green", alpha = 0.2, size = 1)
```


