---
title: "Lab08-Simple-Function"
author: "Betty Cui"
date: "3/14/2018"
output: html_document
---


```{r}
vec <- c(3,1,4)
for (j in vec){
  print(j)
  print(j * 3)
}
```



```{r}
vec2 <- c(3, 1, 4)
vec2 <- rep(0, length(vec))
for (i in 1:3){
  vec2[i] = vec[i] * 3
  print(vec2[i])
}
```


```{r}
## Series 1
n <- 100
summation <- 0

# Do the summation
for(i in 0:n){
  summation <- summation + 0.5 ^ i
}
summation
 # as we can see, when n increases, sum converges to 2

## Series 2
n <- 1000
summation <- 0

# Do the summation
for(i in 0:n){
  summation <- summation + (1/9) ^ i
}
summation
# As we can see, when n increases, series 2 converges to 1.125

```

```{r}
## Arithmetic Sequence
n <- 2
summation3 <- 0

# Do the summation
for(i in 1:n){
  summation3 <- summation3 + 3 + (i - 1) * 3
}
summation3
```

```{r}
## Geometric sequence
n <- 5
summation4 <- 0

# Do the summation
for(i in 1:n){
  summation4 <- summation4 + 3 * 2^(i -1)
}
summation4

```



```{r}
## Sine Approximation
n <- 10
summation5 <- 0
x <- 2

# Calulation the sine function first
sin(x)

# Do the summation
for(i in 1:n){
  summation5 <- summation5 + (-1)^(i+1) * x^(2*i - 1) / factorial(2*i -1)
}
summation5


```



```{r}
## Matrix Loop
set.seed(123)
X <- matrix(rnorm(12), nrow = 4, ncol = 3)
X
```{r}
# Creating a empty matrix Y which is the same size as X
Y <- matrix(NA, nrow = 4, ncol = 3)

# Using the loop to convert the matrix
for (i in 1:nrow(X)) {
  for (j in 1:ncol(X)) {
    if (X[i, j] < 0){
      Y[i, j] <- X[i, j]^2
    }else
    {
      Y[i, j] <- sqrt(X[i, j])
    }
  }
}

```
------------------------------------------------------------------------


```{r}
## Average
average <- function(x){
  i <- 1
  summation6 <-  0
  while (i <= length(x)) {
  summation6 <- summation6 + x[i]
  i <- i + 1
  }
  print(summation6 / length(x))
}

x <- 1:100
average(x)
```

```{r}
## Standard Deviation
stdev <- function(x){
  i <- 1
  summation7 <-  0
  while (i <= length(x)) {
  summation7 <- summation7 + (x[i] - mean(x))^2
  i <- i + 1
  }
  summation7 <- summation7/(length(x) - 1)
  print(sqrt(summation7))
}

x <- 1:100
stdev(x)
```

```{r}
# random distance matrix
num_letters <- length(LETTERS)
set.seed(123)
values <- sample.int(num_letters) 
distances <- values %*% t(values)
diag(distances) <- 0
dimnames(distances) <- list(LETTERS, LETTERS)
distances[1:5, 1:5]
```

```{r}
get_dist <- function(dists, ltrs) {
  if (!all(ltrs %in% LETTERS)) {
    stop('Unrecognized character')
  }
  sum_dist <- 0
  for (l in 1:(length(ltrs)-1)) {
    from <- ltrs[l]
    to <- ltrs[l+1]
    sum_dist <- sum_dist + dists[from,to]
  }
  sum_dist
}
vec <- c('E', 'D', 'A')
get_dist(distances, vec)
error <- c('E', 'D', ')')

```

```{r}
cal <- c('C', 'A', 'L')
get_dist(distances, cal)
stats <- c('S', 'T', 'A', 'T', 'S')
get_dist(distances, stats)
oski <- c('O', 'S', 'K', 'I')
get_dist(distances, oski)
zzz <- rep('Z', 3)
get_dist(distances, zzz)
lets <- LETTERS
get_dist(distances, lets)
first = c('P', 'I', 'N', 'R', 'U', 'I')
get_dist(distances, first)
last = c('Y', 'U')
get_dist(distances, last)

## Geometric mean
# Continue the loop as long as x is not finished
gmean <- function(x){
  i <- 1
  product <-  1
  while (i <= length(x)) {
  product <- product * x[i]
  i <- i + 1
  }
  product <-  product^(1/length(x))
  print(product)
}

x <- 1:50
gmean(x)

```

```{r}
strings <- list(
  cal = cal,
  stats = stats,
  oski = oski,
  zzz=zzz,
  lets = lets,
  first = first,
  last = last
)

string_dists <- rep(0, length(strings) )
for (i in 1:length(strings)){
  string_dists[i] <- get_dist(distances, strings[[i]])
}
string_dists
```