---
title: "Lab04-Betty-Cui"
author: "Betty Cui"
date: "2/14/2018"
output: html_document
---
# The character delimiter is comma.
# On this first dataset, there is not such a row for column names. The second one has a row for column names.
# There are no missing values.
# They are numeric and character.
```{eva = False}
# The abalone.data file is ASCII text.
# Totally I have 4177 newline count, 4177 word count, and 191873 byte count.
# To only obtain the number of lines in abalone.data, we should choose wc -l, --lines.
# The first 10 lines of abalone file are M,0.455,0.365,0.095,0.514,0.2245,0.101,0.15,15
M,0.35,0.265,0.09,0.2255,0.0995,0.0485,0.07,7
F,0.53,0.42,0.135,0.677,0.2565,0.1415,0.21,9
M,0.44,0.365,0.125,0.516,0.2155,0.114,0.155,10
I,0.33,0.255,0.08,0.205,0.0895,0.0395,0.055,7
I,0.425,0.3,0.095,0.3515,0.141,0.0775,0.12,8
F,0.53,0.415,0.15,0.7775,0.237,0.1415,0.33,20
F,0.545,0.425,0.125,0.768,0.294,0.1495,0.26,16
M,0.475,0.37,0.125,0.5095,0.2165,0.1125,0.165,9
F,0.55,0.44,0.15,0.8945,0.3145,0.151,0.32,19
# In order to onlt display the first 5 lines of abalone file, we should us "head -n5 abalone.data"
# The last 10 lines of abalone.data are
M,0.5,0.38,0.125,0.577,0.269,0.1265,0.1535,9
F,0.515,0.4,0.125,0.615,0.2865,0.123,0.1765,8
M,0.52,0.385,0.165,0.791,0.375,0.18,0.1815,10
M,0.55,0.43,0.13,0.8395,0.3155,0.1955,0.2405,10
M,0.56,0.43,0.155,0.8675,0.4,0.172,0.229,8
F,0.565,0.45,0.165,0.887,0.37,0.239,0.249,11
M,0.59,0.44,0.135,0.966,0.439,0.2145,0.2605,10
M,0.6,0.475,0.205,1.176,0.5255,0.2875,0.308,9
F,0.625,0.485,0.15,1.0945,0.531,0.261,0.296,10
M,0.71,0.555,0.195,1.9485,0.9455,0.3765,0.495,12
# In order to get only the last three lines, we should use tail -n3 abalone.data 
```

```{r}
abalone <- read.table("abalone.data", sep = ",")
head(abalone)
tail(abalone)
str(abalone, vec.len = 1)
```

```{r}
# My Turn
read.csv("abalone.data")
column_names <- c(
    'sex',
    'length',
    'diameter',
    'height',
    'whole_weight',
    'shucked_weight',
    'viscera_weight',
    'shell_weight',
    'rings'
)
column_types <- c(
    'character',
    'real',
    'real',
    'real',
    'real',
    'real',
    'real',
    'real',
    'integer'   
)
abalone <- read.table(
    'abalone.data',
    col.names = column_names,
    colClasses = column_types,
    sep = ","
)
```

```{r}
read.csv("bridges.data.version1")
read.table("bridges.data.version1")
str("bridges.data.version1")
summary("bridges.data.version1")
head("bridges.data.version1")
tail("bridges.data.version1")
dim("bridges.data.version1")
names("bridges.data.version1")
colnames("bridges.data.version1")
nrow("bridges.data.version1")
ncol("bridges.data.version1")
```

#NBA Players Data Frame
```{r}
#Creating vectors
  Player = c("Thompson", "Curry", "Green", "Durant", "Pachulia")
  Position = c("SG", "PG", "PF", "SF", "C")
  Salary = c(16663575, 12112359, 15330435, 26540100, 2898000)
  Points = c(1742, 1999, 776, 1555, 426)
  PPG = c(22.3, 25.3, 10.2, 25.1, 6.1)
  Rookie = c(FALSE, FALSE, FALSE, FALSE, FALSE)

# Create a table using the data frame directly
GSW1 <- data.frame(Player, Position, Salary, Points, PPG, Rookie)

GSW1

# Create a table using the list first
GSWlist <- list(
  Player = c("Thompson", "Curry", "Green", "Durant", "Pachulia"),
  Position = c("SG", "PG", "PF", "SF", "C"),
  Salary = c(16663575, 12112359, 15330435, 26540100, 2898000),
  Points = c(1742, 1999, 776, 1555, 426),
  PPG = c(22.3, 25.3, 10.2, 25.1, 6.1),
  Rookie = c(FALSE, FALSE, FALSE, FALSE, FALSE)
)

GSW2 <- data.frame(GSWlist)

GSW2

# Change the factor to characters
GSW1$Player <- as.character(GSW1$Player)
GSW1$Points <- as.integer(GSW1$Points)
# Get the data structure
str(GSW1)

# Combine the vectors
GSW3 <- cbind(Player, Position, Salary, Points, PPG, Rookie, deparse.level = 1)

# Convert the cbind object to a data frame
GSW3 <- data.frame(GSW3)

GSW3
```

```{r}

csv <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/data/nba2017-players.csv"
download.file(url = csv, destfile = 'nba2017-players.csv', method = 'curl')
dat <- read.csv('nba2017-players.csv', stringsAsFactors = FALSE)
```

A first check-up is to examine the dimensions of the data frame with `dim()`:

```{r}
# dimensions (# of rows, # of columns)
dim(dat)
```

```{r}
# display first few rows
head(dat)
```

For a more detailed description of how R is treating the data type in each 
column, you should use the structure function `str()`.

```{r}
# check the structure
str(dat, vec.len = 1)
```

- Display the last 5 rows of the data.

```{r}
#display last 5 rows
dat[436:441, ]
```

- Display those rows associated to players having height less than 70 inches tall.

```{r}
#display the height less than 70
dat[dat$height < 70, ]
```

- Of those players that are centers (position `C`), display their names and salaries.

```{r}
#display the names and salaries of the centers
dat[dat$position == "C", c("player", "salary")]
```

- Create a data frame `durant` with Kevin Durant's information (i.e. row).

```{r}
#finding kevin durant
durant <- data.frame(
  dat[dat$player == "Kevin Durant", ]
)
durant
```

- Create a data frame `ucla` with the data of players from college UCLA ("University of California, Los Angeles").

```{r}
#finding ucla players
ucla <- data.frame(
  dat[dat$college == "University of California, Los Angeles", ]
)
ucla
```

- Create a data frame `rookies` with those players with 0 years of experience. 

```{r}
#finding rookies
rookies <- data.frame(
  dat[dat$experience == 0, ]
)
rookies
```

- Create a data frame `rookie_centers` with the data of Center rookie players. 

```{r}
#Finding center in the rookies
rookie_centers <- data.frame(
  dat[(dat$experience == 0 & dat$position == "C"), ]
)
rookie_centers
```

- Create a data frame `top_players` for players with more than 50 games and more than 100 minutes played.

```{r}
#Finding top players
top_players <- data.frame(
  dat[(dat$games >= 50 & dat$minutes >= 100), ]
)
top_players
```

- What's the largest height value?

```{r}
#Finding larhtgerst height
max(dat$height)
```

- What's the minimum height value?

```{r}
# get the minimum value of height
min(dat$height)
```

- What's the overall average height?

```{r}
#get the mean value of the height
mean(dat$height)
```

- Who is the tallest player?

```{r}
#get the player that has the max height
dat$player[dat$height == max(dat$height)]
```

- Who is the shortest player?

```{r}
#get the shortest player
dat$player[dat$height == min(dat$height)]
```

- Which are the unique teams?

```{r}
unique(dat$team)
```

- How many different teams?

```{r}
#get total number of teams
length(unique(dat$team))
```

- Who is the oldest player?
```{r}
#get the oldest player
dat$player[dat$age == max(dat$age)]
```

- What is the median salary of all players?

```{r}
#get the median salary
median(dat$salary, na.rm = FALSE)
```

- What is the median salary of the players with 10 years of experience or more?
```{r}
#get the median of the salary of 10-year player
median(dat$salary[dat$experience >= 10], na.rm = FALSE)
```

- What is the median salary of Shooting Guards (SG) and Point Guards (PG)?

```{r}
#get the median of the shooting guards and points guards
median(dat$salary[dat$position == "SG" | dat$position == "PG"], na.rm = FALSE)
```

- What is the median salary of Power Forwards (PF), 29 years or older, and 74 inches tall or less?

```{r}
median(dat$salary[(dat$position == "PF" & dat$age >= 29) & dat$height <= 74], na.rm = FALSE)
```

- How many players scored 4 points or less?
```{r}
#get the player scored 4 points or less
length(dat$player[dat$points <= 4])
```

- Who are those players who scored 4 points or less?

```{r}
#name of the players who scored 4 points less
dat$player[dat$points <= 4]
```

- Who is the player with 0 points?

```{r}
#player with 0 points
dat$player[dat$points == 0]
```

- How many players are from "University of California, Berkeley"? 

```{r}
#players from UCB
dat[dat$college == "University of California, Berkeley", "player"]
```

- Are there any players from "University of Notre Dame"? If so how many and who are they?

```{r}
#get the player from university of notre dame
dat$player[dat$college == "University of Notre Dame"]
length(dat$player[dat$college == "University of Notre Dame"])
```

- Are there any players with weight greater than 260 pounds? If so how many and who are they?

```{r}
#get the player greater than 260 pounds
dat$player[dat$weight > 260]
length(dat$player[dat$weight > 260])
```

- How many players did not attend a college in the US?

```{r}
# get the number of the players who did not attend a college
length(dat[dat$college == "", ])
```

- Who is the player with the maximum rate of points per minute?
```{r}
# get player with max rate of points per minute
dat$player[dat$points/dat$minutes == max(dat$points/dat$minutes)]
```

- Who is the player with the maximum rate of three-points per minute?
```{r}
# get max rate of three-point
dat$player[dat$points3/dat$minutes == max(dat$points3/dat$minutes)]
```

- Who is the player with the maximum rate of two-points per minute?
```{r}
#get max 2 point player
dat$player[dat$points2/dat$minutes == max(dat$points2/dat$minutes)]
```

- Who is the player with the maximum rate of one-points (free-throws) per minute?
```{r}
#get max 1 point player
dat$player[dat$points1/dat$minutes == max(dat$points1/dat$minutes)]
```

- Create a data frame `gsw` with the name, height, weight of Golden State Warriors (GSW)
```{r}
#create a data frame of gsw
gsw <- data.frame(
  name = dat$player[dat$team == "GSW"],
  height = dat$height[dat$team == "GSW"],
  weight = dat$weight[dat$team == "GSW"]
)

gsw
```

- Display the data in `gsw` sorted by height in increasing order (hint: see `?sort` and `?order`)
```{r}
#incresing order of height
gsw <- gsw[order(gsw$height, na.last = TRUE, decreasing = FALSE, method = c("auto")), ]
gsw
```

- Display the data in gsw by weight in decreasing order (hint: see `?sort` and `?order`)
```{r}
#decreasing order of weight
gsw <- gsw[order(gsw$weight, na.last = TRUE, decreasing = TRUE, method = c("auto")), ]
gsw
```

- Display the player name, team, and salary, of the top 5 highest-paid players (hint: see `?sort` and `?order`)
```{r}
#display the highest 5 salaries
highest_paid <- dat[order(dat$salary, na.last = TRUE, decreasing = TRUE, method = "auto"), c("player", "team", "salary")]

highest_paid[1:5, ]
```

- Display the player name, team, and points3, of the top 10 three-point players (hint: see `?sort` and `?order`)
```{r}
#display the top 10 3-pointers
top_threepoint <- dat[order(dat$points3, na.last = TRUE, decreasing = TRUE, method = "auto"), c("player", "team", "points3")]

top_threepoint[1:10, ]
```

## Group By

Group-by operations are very common in data analytics. Without dedicated 
functions, these operations tend to be very hard (labor intensive). 

__Quick try__: Using just bracket notation, try to create a data frame with 
two columns: the team name, and the team payroll (addition of all player salaries).

So what functions can you use in R to perform group by operations? In base R, 
the main function for group-by operations is `aggregate()`.

Here's an example using `aggregate()` to get the median salary, grouped by team:

```{r}
aggregate(dat$salary, by = list(dat$team), FUN = median)
```

The same example above can also be obtained with `aggreagte()` using formula 
notation like this:

```{r}
aggregate(salary ~ team, data = dat, FUN = median)
```


Here's another example using `aggregate()` to get the average height and 
average weight, grouped by position:

```{r}
aggregate(dat[ ,c('height', 'weight')], by = list(dat$position), FUN = mean)
```

The same example above can also be obtained with `aggreagte()` using formula 
notation like this:

```{r}
aggregate(. ~ position, data = dat[ ,c('position', 'height', 'weight')],
          FUN = mean)
```

### Your turn

- Create a data frame with the average height, average weight, and average 
age, grouped by position
```{r}
aggregate(.~ dat$position, data = dat[ , c("height", "weight", "age")], FUN = mean)
```

- Create a data frame with the average height, average weight, and average 
age, grouped by team
```{r}
aggregate(.~ dat$team, data = dat[ , c("height", "weight", "age")], FUN = mean)
```

- Create a data frame with the average height, average weight, and average 
age, grouped by team and position.
```{r}
aggregate(dat[ , c("height", "weight", "age")], by = list(dat$team, dat$position), FUN = mean)
```

- Difficult: Create a data frame with the minimum salary, median salary,
mean salary, and maximum salary, grouped by team and position.
```{r}
table1 <- aggregate(dat[ , "salary"], by = list(dat$team, dat$position), FUN = min)
table2 <- aggregate(dat[ , "salary"], by = list(dat$team, dat$position), FUN = median)
table3 <- aggregate(dat[ , "salary"], by = list(dat$team, dat$position), FUN = mean)
table4 <- aggregate(dat[ , "salary"], by = list(dat$team, dat$position), FUN = max)

salary_table <- cbind(table1, table2[ , 3], table3[ , 3], table4[ , 3])
attributes(salary_table)$names[3] <- "min_salary"
attributes(salary_table)$names[4] <- "median_salary"
attributes(salary_table)$names[5] <- "mean_salary"
attributes(salary_table)$names[6] <- "max_salary"

salary_table
```
