---
title: "hw04"
author: "Betty Cui"
date: "4/12/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
library(stringr)
library(ggplot2)
library(readr)
```

```{r}
source("/Users/cj/Desktop/stats133/hw04/code/archive-functions.R")
```

```{r}
raw_data <- read_archive('stringr')
clean_data <- clean_archive(raw_data)
write.csv(clean_data, file = "/Users/cj/Desktop/stats133/hw04/data/stringr-archive.csv", row.names = FALSE)
plot_archive(clean_data)
```

```{r}
dat1 <- read_archive("dplyr")
dat1 <- read_archive("dplyr")
rex1 <- clean_archive(dat1)
write.csv(rex1, file = "/Users/cj/Desktop/stats133/hw04/data/dplyr-archive.csv", row.names = FALSE)
#plot_archive(rex1)
```

```{r}
dat2 <- read_archive("XML")
rex2 <- clean_archive(dat2)
write.csv(rex2, file = "/Users/cj/Desktop/stats133/hw04/data/xml-archive.csv", row.names = FALSE)
```

```{r}
dat3 <- read_archive("ggplot2")
rex3 <- clean_archive(dat3)
write.csv(rex3, file = "/Users/cj/Desktop/stats133/hw04/data/ggplot2-archive.csv", row.names = FALSE)
```

```{r}
dat4 <- read_archive("knitr")
rex4 <- clean_archive(dat4)
write.csv(rex4, file = "/Users/cj/Desktop/stats133/hw04/data/knitr-archive.csv", row.names = FALSE)
```

```{r}
dat <- rbind(rex1, rex2, rex3, rex4)
```

```{r}
picture1 <- ggplot(data = dat, aes(x = date, y = size, color = name)) +
    geom_step() +
    ggtitle('Four package: timeline of version sizes') +
    labs(y = 'Size (Kilobytes)') 
picture1
```

```{r}
picture2 <- ggplot(data = dat, aes(x = date, y = size, color = name)) +
    geom_step() +
    ggtitle('Four package: timeline of version sizes') +
    labs(y = 'Size (Kilobytes)') +
    facet_wrap(~name, scales = 'free')
picture2
```

```{r}
### 3.1 Number of the characters in the tweet contents

dat <- read.csv2("/Users/cj/Desktop/stats133/hw04/data/text-emotion.csv")
num <- rep(0, length(dat[ ,1]))
for (i in 1:length(dat[, 1])) {
  num[i] <- nchar(as.character(dat[i, 1]))
}
hist(num, xlab="Character Length", breaks=seq(0,210,l=40))

### 3.2 Number of Mentions


# 3.2
mentions <- rep(0, length(dat[ ,1]))
value = integer(0)
for (i in 1:length(dat[, 1])) {
  # convert words to vector, get rid of space, "\"", and ","
  ch <- as.character(dat[i, 1])
  h <- str_split(ch, ",")
  h <- str_split(h, " ")
  h <- str_split(h, "\"")
  h <- as.vector(unlist(h))
  # if h contains @
  m <- grep('@', h)
  if (!identical(m, integer(0))) {
    mentions[i] = length(m)
  }
}
barplot(mentions)
# Display the content with 10 tags
for (i in 1:length(mentions)) {
  if (mentions[i] == 10) {
    print(dat[i, 1])
  }
}



### 3.3
contents <- data.frame(read.csv2("/Users/cj/Desktop/stats133/hw04/data/contents.csv"))
string_tweet <- data.frame(read.csv2("/Users/cj/Desktop/stats133/hw04/data/contents.csv"))



count_hash <- function(paulo){
  return(str_count(paulo, pattern = "#" ))
}
#[A-z_][\\w]+
count_hash(contents[ , 1])

sum(count_hash(contents[ , 1])) #the number of hashtags in the tweet contents
table(count_hash(contents[ , 1]))

barplot(table(count_hash(contents)))

sex <- str_extract(string_tweet,pattern = "#[A-z][\\w]+")
sex <- sex[!is.na(sex)]

#average length
mean(nchar(sex)-1) #7.775068

tab_hash <- table(nchar(sex)-1)
names(tab_hash)[which(tab_hash==max(tab_hash))]

## Most common length is 4




